---
title: "Project 1"
author: "Jordan Tanley"
date: "6/8/2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(knitr)
knitr::opts_chunk$set(fig.path = "~/images/")
```


# Packages

The required packages to use the following functions and summary statistics are as follows. 

* [`tidyverse`](https://www.tidyverse.org/): Used for data handling  
* [`dplyr`](https://dplyr.tidyverse.org/): Helpful for data manipulation  
* [`httr`](https://httr.r-lib.org/): Used to access APIs  
* [`jsonlite`](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html): Aslo used to access APIs  
* [`ggplot2`](https://ggplot2.tidyverse.org/): Used for creating plots


```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(httr)
library(jsonlite)
library(ggplot2)
```


# API Functions

Here, I will define the function I created to access information from the PokeAPI. 

### `pokemonfacts`

This function has one required input:

* `pokemonName`: The user has a choice between 10 Pokemon. These include Weedle, Butterfree, Venusaur, Bulbasaur, Squirtle, Pidgeotto, Wartortle, Blastoise, Beedrill or Charmeleon. The name must be entered as a string, (i.e. "weedle"). Capitolization does not matter here.  

And several optional inputs:

* `base_exp`: When `TRUE`, this option provides the base experience for the selected Pokemon. When `FALSE`, the value is `NA`.  

* `height`: When `TRUE`, this option provides the height for the selected Pokemon. When `FALSE`, the value is `NA`.  

* `weight`: When `TRUE`, this option provides the weight for the selected Pokemon. When `FALSE`, the value is `NA`.  

* `type`: When `TRUE`, this option provides the number of types the selected Pokemon is classified as. When `FALSE`, the value is `NA`.  

* `color`: When `TRUE`, this option provides the color for the selected Pokemon. When `FALSE`, the value is `NA`.  

* `shape`: When `TRUE`, this option provides the shape for the selected Pokemon. When `FALSE`, the value is `NA`.  

When run, this function provides the selected information in the form of a data set with one row and seven columns. These collumns correspond to the name and all 6 options. 

```{r}
pokemonfacts <- function(pokemonName, base_exp = TRUE, height = TRUE, weight = TRUE, type = TRUE, color = TRUE, shape = TRUE){
  
  baseURL <- "https://pokeapi.co/api/v2/"
  output <- data.frame(matrix(nrow = 1))
  
  if (tolower(pokemonName) %in% "charmeleon") {
    endURL1 <- "pokemon/charmeleon"
    output$name <- "charmeleon"
    col <- GET("https://pokeapi.co/api/v2/pokemon-species/charmeleon/")
  }
  
  else if (tolower(pokemonName) %in% "pidgeotto") {
    endURL1 <- "pokemon/pidgeotto"
    output$name <- "pidgeotto"
    col <- GET("https://pokeapi.co/api/v2/pokemon-species/pidgeotto/")
  }
  
  else if (tolower(pokemonName) %in% "squirtle") {
    endURL1 <- "pokemon/squirtle"
    output$name <- "squirtle"
    col <- GET("https://pokeapi.co/api/v2/pokemon-species/squirtle/")
  }
  
  else if (tolower(pokemonName) %in% "bulbasaur") {
    endURL1 <- "pokemon/bulbasaur"
    output$name <- "bulbasaur"
    col <- GET("https://pokeapi.co/api/v2/pokemon-species/bulbasaur/")
  }
  
  
  else if (tolower(pokemonName) %in% "venusaur") {
    endURL1 <- "pokemon/venusaur"
    output$name <- "venusaur"
    col <- GET("https://pokeapi.co/api/v2/pokemon-species/venusaur/")
  }
  
  else if (tolower(pokemonName) %in% "butterfree") {
    endURL1 <- "pokemon/butterfree"
    output$name <- "butterfree"
    col <- GET("https://pokeapi.co/api/v2/pokemon-species/butterfree/")
  }
  
  else if (tolower(pokemonName) %in% "wartortle") {
    endURL1 <- "pokemon/wartortle"
    output$name <- "wartortle"
    col <- GET("https://pokeapi.co/api/v2/pokemon-species/wartortle/")
  }
  
  else if (tolower(pokemonName) %in% "weedle") {
    endURL1 <- "pokemon/weedle"
    output$name <- "weedle"
    col <- GET("https://pokeapi.co/api/v2/pokemon-species/weedle/")
  }
  
  else if (tolower(pokemonName) %in% "blastoise") {
    endURL1 <- "pokemon/blastoise"
    output$name <- "blastoise"
    col <- GET("https://pokeapi.co/api/v2/pokemon-species/blastoise/")
  }
  
  else if (tolower(pokemonName) %in% "beedrill") {
    endURL1 <- "pokemon/beedrill"
    output$name <- "beedrill"
    col <- GET("https://pokeapi.co/api/v2/pokemon-species/beedrill/")
  }
  
  else {stop("Not one of the pokemon name options. 
             Choose one of: Weedle, Butterfree, Venusaur, Bulbasaur, Squirtle, Pidgeotto, Wartortle, Blastoise, Beedrill or Charmeleon")}
  
    poke <- paste0(baseURL, endURL1) %>% GET() 
    parsed <- poke$content %>% rawToChar() %>% fromJSON()  
  
    if (base_exp) output$base_experience <- parsed$base_experience
    else {output$base_experience <- NA}
    
    if (height)   output$height <- parsed$height 
    else {output$height <- NA}
    
    if (weight) output$weight <- parsed$weight
    else {output$weight <- NA}
    
    if (type) output$types <- length(parsed$types$type$name)
    else {output$types <- NA}
    
    if (color || shape) {
      
      parsed2 <- col$content %>% rawToChar() %>% fromJSON()
      
      if (color) {output$color <- parsed2$color$name}         
      else {output$color <- NA}   
      
      if (shape) {output$shape <- parsed2$shape$name}
      else {output$shape <- NA}
      
    }
  
  Out <- subset(output, select = -1)
  
  return(Out)
}
```


### `cleaning`

This function can be used to drop any missing values in the outputted dataframe that the previous function creates. The cases where this is useful are when the user wishes to only see a subset of the variables offered. 


```{r}
cleaning <- function(data){
  data %>%
    select_if(~ !any(is.na(.)))
}
```

Here is an example of this function at work. Originally, the dataset `c` has all 7 variables, with `shape` and `color` as `NA`. After using the `cleaning()` function, the `NA`s are dropped and the resulting dataset has 5 variables. 

```{r}
c <- pokemonfacts("charmeleon", shape = F, color = F)
cleaning(c)
```


### `pokemon`

This function is used to combine several datasets that were created using the `pokemonfacts()` function. This allows the user to create a master list, or a Pokedex, of all of the pokemon that they choose to investigate. Implementation is the same as `rbind()`, but the name is changed to match the theme. 

```{r}
pokemon <- function(pokemon1, pokemon2, ...){
  out <- rbind(pokemon1, pokemon2, ...)
  return(out)
}
```


And that is all of the functions I created to access and use this Pokemon API! Next, I will begin some exploratory data analysis. 




# Exploratory Data Analysis

Next, let's use some of the API endpoints to pull data into a dataset. 

## Data

The first step is to use the `pokemonfacts()` function to access several different endpoints. Here, I will use 10 separate endpoints, leading to 10 separate dataframes - one for each pokemon. Next, I will use the `pokemon()` function to concatinate the 10 datasets into one long dataset with 7 variables and 10 rows. Then, using piping from the `dplyr` package, I will convert the resulting dataframe to a tibble. Tibbles are helpful as they have better printing qualities than a dataframe. Finally, the resulting tibble is saved as an R object called `pokedex`. 

```{r}
# access API
c <- pokemonfacts("charmeleon")
p <- pokemonfacts("pidgeotto")
s <- pokemonfacts("squirtle")
b <- pokemonfacts("bulbasaur")
v <- pokemonfacts("venusaur")
f <- pokemonfacts("butterfree")
w <- pokemonfacts("weedle")
bl <- pokemonfacts("blastoise")
wa <- pokemonfacts("wartortle")
bee <- pokemonfacts("beedrill")
# use pokemon() function to combine all datasets into one larger one. Convert that dataframe to a tibble.
pokedex <- pokemon(s, b, p, c, v, f, w, bl, wa, bee) %>% as_tibble()
# view pokedex tibble
pokedex
```

For further analysis, lets create a new variable based on the existing variables. I am interested in examining how weight classes effect their shapes, number of types, and other areas. With this in mind, lets create a `size` variable that classifies `weight` into bins of "Small", "Medium", or "Large". Because of the range of weight values, I used a wider range of values for "Medium" than for the other two bins. First, lets

```{r}
cutoff <- (max(pokedex$weight) - min(pokedex$weight))/4

pokedex$size <- ifelse(pokedex$weight < cutoff, "Small", 
                  ifelse(pokedex$weight < 2*cutoff, "Medium", 
                    ifelse(pokedex$weight >= 3*cutoff,  "Large",  NA)))

pokedex
```


## Tables

You should create some contingency tables

```{r}
table(pokedex$color)
table(pokedex$shape)
table(pokedex$color, pokedex$shape)
table(pokedex$size, pokedex$shape)
```


## Numerical Summaries

You should create numerical summaries for some quantitative variables at each setting of some of your categorical variables

```{r}
pokedex %>% group_by(color) %>%
  summarize(min = min(weight), avg = mean(weight), med = median(weight), max = max(weight), var = var(weight))
```

```{r}
pokedex %>% group_by(shape) %>%
  summarize(min = min(weight), avg = mean(weight), med = median(weight), max = max(weight), var = var(weight))
```






## Graphical Summaries 

You should create at least five plots utilizing coloring, grouping, etc. All plots should have nice labels and titles.

You should have at least one bar plot, one histogram, one box plot, and one scatter plot


```{r}
g <- ggplot(data = pokedex, aes(x = shape, fill = as.factor(types)))
g + geom_bar()   + labs(x = "Pokemon Shape", y = "Frequency", title = "Bar Plot of Pokemon Shapes") + scale_fill_brewer(palette = "Accent") +  guides(fill=guide_legend(title="Number of Types") ) + theme_classic()
```


```{r}
g <- ggplot(data = pokedex, aes(x = weight))
g + geom_histogram(binwidth = 50, fill = "#c2add7") + labs(x = "Pokemon Weight", y = "Frequency", title = "Histogram of Pokemon Weights") + theme_classic()
```

```{r}
g <- ggplot(pokedex, aes(x = weight, y = height))
g + geom_point(aes(color = shape))  + labs(x = "Pokemon Weight", y = "Pokemon Height", title = "Scatterplot of Pokemon Weights vs Heights") + scale_color_brewer(palette = "Accent") +  guides(color=guide_legend(title="Shape")) + theme_classic() 
```


```{r}
# set up plot
g <- ggplot(pokedex, aes(x = shape, y = weight))
# white boxplots with scatterplot over them, scatter is grouped and colored by shape 
g + geom_boxplot(fill = "white") + geom_jitter(aes(group = shape, col = shape)) + labs(title = "Boxplot for Weight for Each Shape") + theme_classic() + scale_color_brewer(palette = "Accent")
```


```{r}
# set up plot
g <- ggplot(pokedex, aes(x = size, y = weight))
# white boxplots with scatterplot over them, scatter is grouped and colored by size 
g + geom_boxplot(fill = "white") + geom_jitter(aes(group = size, col = size)) + labs(title = "Boxplot for Weight for Each Shape") + theme_classic() + scale_color_brewer(palette = "Accent")
```


```{r}

```












